Communication complexity: Chapter 6 \\ One-way communication and simultaneous message passing

### 2019-04-18 CS860-CommunicationComplexity/assignments/Chapter6.tex
Correcting Chapter 2
SMP equality

\begin{proof}
Consider a one-way randomized communication protocol for $\Index$ with success probability $\frac{2}{3}$, achieving communication cost $R^\rightarrow(\Index)$. Let $X$ and $Y$ be random variables for Alice and Bob's inputs respectively, each distributed uniformly and independently at random. Let $M$ be a random variable for the message $m$ that Alice sends Bob during the operation of this protocol. Our goal is to use information-theoretic techniques to lower bound $|M|$, the maximum length of a message in $M$'s support. We have,
\begin{align*}
|M| &\geq H(M)\\
&\geq I(X;M) \\
&=\sum_{i=1}^n I(X_i; M \mid X_1\dots X_{i-1}) &\text{by chain rule} \\
&=\sum_{i=1}^n H(X_i\mid X_1\dots X_{i-1}) - H(X_i\mid M, X_1\dots X_{i-1})\\
&= n - \sum_{i=1}^n H(X_i\mid M, X_1\dots X_{i-1}) &\text{since the bits of $X$ are chosen uniformly independently} \\
&\geq n-\sum_{i=1}^n H(X_i\mid M).
\end{align*} 
We claim $H(X_i\mid M)\leq \frac{2}{3}$. With this claim in hand, we see that
$$|M| \geq n-\frac{2}{3}n =\Omega(n).$$
To see the claim, we use that the protocol succeeds with probability $\frac{2}{3}$. We expand the conditional entropy as an expectation and compute
$$H(X_i\mid M) = \sum_{m\in M} H(X_i\mid M=m)\cdot \Pr(M=m) = \sum_{m\in M}(\frac{2}{3}\log(\frac{3}{2}) + \frac{1}{3}\log(3))\Pr(M=m) \leq \frac{2}{3}.$$
The second inequality follows by the success probability of the protocol. The probability that the protocol say $X_i$ is the correct bit is $\frac{2}{3}$, and the probability that $X_i$ is the wrong bit is $\frac{1}{3}$.
\end{proof}

@@

	We want to show that for each $x \in \calX$ there exists a set of messages $S_x \subseteq M_x$ such that for any $y \in \calY$, and message $m_y \in M_y$, the protocol correctly outputs $f(x,y)$ on the majority of messages $S_x$ with $m_y$. Let $\tau(m_x,m_y)$ be the referees (deterministic) output on messages $(m_x,m_y)$ for protocol $\pi$. I.e.
	$$\frac{2}{3} \leq \Pr(\pi(x,y) = f(x,y)) = \Pr_{m_x,m_y}(\tau(m_x,my) = f(x,y)).$$
	Fix $y\in \calY$, and $m_y \in M_y$. Drawn a uniform random subset $S_x \subseteq M_x$  of size $t$. By a Chernoff Bound, the probability that $\tau(m_x,m_y) \neq f(x,y)$ for at least half of message pairs $m_x, m_y$ is at most
	$$2^{-\Omega(t)}.$$
	Since $\pi$ has cost $R^{\parallel,\mathrm{priv}}(f)$, over all $y$ there are at most $2^{R^{\parallel,\mathrm{priv}}(f)}$ messages the referee sees from Bob. So by union bound, the probability $\tau(m_x,m_y) \neq f(x,y)$ for some message $m_y$ is at most
	$$2^{R^{\parallel,\mathrm{priv}}(f)}2^{-\Omega(t)}.$$
	This probability is strictly less than one for $t = \Omega(R^{\parallel,\mathrm{priv}}(f))$. Thus by the probabilistic method the set $S_x$ exists. By a symmetric argument, the analogous set $S_y$ exists for each $y \in \calY$.

@@

@@

\begin{proof}
\replacethistext{Enter your proof here.}
\end{proof}

>>>>

\begin{proof}
Consider a one-way randomized communication protocol for $\Index$ with success probability $\frac{2}{3}$, achieving communication cost $R^\rightarrow(\Index)$. Let $X$ and $Y$ be random variables for Alice and Bob's inputs respectively, each distributed uniformly and independently at random. Let $M$ be a random variable for the message $m$ that Alice sends Bob during the operation of this protocol. Our goal is to use information-theoretic techniques to lower bound $|M|$, the maximum length of a message in $M$'s support. We have,
\begin{align*}
|M| &\geq H(M)\\
&\geq I(X;M) \\
&=\sum_{i=1}^n I(X_i; M \mid X_1\dots X_{i-1}) &\text{by chain rule} \\
&=\sum_{i=1}^n H(X_i\mid X_1\dots X_{i-1}) - H(X_i\mid M, X_1\dots X_{i-1})\\
&= n - \sum_{i=1}^n H(X_i\mid M, X_1\dots X_{i-1}) &\text{since the bits of $X$ are chosen uniformly independently} \\
&\geq n-\sum_{i=1}^n H(X_i\mid M).
\end{align*} 
We claim $H(X_i\mid M)\leq \frac{2}{3}$. With this claim in hand, we see that
$$|M| \geq n-\frac{2}{3}n =\Omega(n).$$
To see the claim, we use that the protocol succeeds with probability $\frac{2}{3}$. We expand the conditional entropy as an expectation and compute
\begin{align*}
	H(X_i\mid M) &= \sum_{m \in M} \Pr(M=m)\mathbb{E}[-\log(\Pr(X_i\mid M=m))]\\
	&\leq \mathbb{E}[-\log(\Pr(X_i\mid M=m)\Pr(M=m)] &\text{ via Jensen's inequality}\\
	&= \mathbb{E}[H(X_i\mid E[M])]\\
	&\leq \frac{2}{3}\log\left(\frac{3}{2}\right) + \frac{1}{3}\log(3)\\
	&\leq \frac{2}{3}.
\end{align*}
The second inequality follows by the success probability of the protocol. The probability that the protocol say $X_i$ is the correct bit is $\frac{2}{3}$, and the probability that $X_i$ is the wrong bit is $\frac{1}{3}$.
\end{proof}

@@

	We want to show that for each $x \in \calX$ there exists a set of messages $S_x \subseteq M_x$ such that for any $y \in \calY$, and message $m_y \in M_y$, the protocol correctly outputs $f(x,y)$ on the majority of messages $S_x$ with $m_y$. Let $\tau(m_x,m_y)$ be the referees (deterministic) output on messages $(m_x,m_y)$ for protocol $\pi$. I.e.
	$$\frac{2}{3} \leq \Pr(\pi(x,y) = f(x,y)) = \Pr_{m_x,m_y}(\tau(m_x,my) = f(x,y)).$$
	Fix $y\in \calY$, and $m_y \in M_y$. Draw a uniform random subset $S_x \subseteq M_x$  of size $t$. By a Chernoff Bound, the probability that $\tau(m_x,m_y) \neq f(x,y)$ for at least half of message pairs $m_x, m_y$ is at most
	$$2^{-\Omega(t)}.$$
	Since $\pi$ has cost $R^{\parallel,\mathrm{priv}}(f)$, over all $y$ there are at most $2^{R^{\parallel,\mathrm{priv}}(f)}$ messages the referee sees from Bob. So by union bound, the probability $\tau(m_x,m_y) \neq f(x,y)$ for some message $m_y$ is at most
	$$2^{R^{\parallel,\mathrm{priv}}(f)}2^{-\Omega(t)}.$$
	This probability is strictly less than one for $t = \Omega(R^{\parallel,\mathrm{priv}}(f))$. Thus by the probabilistic method the set $S_x$ exists. By a symmetric argument, the analogous set $S_y$ exists for each $y \in \calY$.

@@

\begin{definition}
	A set $M \subseteq \{0,1\}^n$ is called a $(n,k,d)$-code if
	\begin{itemize}
		\item $|M| = 2^k$, and 
		\item For every pair $x\neq y \in M$, the Hamming distance between $x$ and $y$ is at least $d$.
	\end{itemize}
\end{definition}

\begin{lemma}\label{lemma:code}
There exists a $(3m,m,m/2)$-code for any $m$.
\end{lemma}

\begin{proof}
We construct the code, greedily. Begin with $M= \emptyset$ and at each iteration choose $x\in \{0,1\}^{3m}\backslash M$ such that the Hamming distance from $x$ to every codeword $y \in M$ is at least $m/2$. Add $x$ to $M$. Terminate when there are no more $x$ possible to add.

We claim that $|M| \geq 2^m$ upon termination of the above procedure. To see this, we bound the number of strings in $\{0,1\}^{3m}$ that any $x \in \{0,1\}^{3m}$ prevents from being chosen in future iterations, i.e. we bound the number of $y \in \{0,1\}^{3m}$ at Hamming distance at most $\frac{m}{2}-1$ from $x$.

There are at most 
$$\sum_{i=1}^{n/2 - 1} {3m \choose i} \leq {3m \choose m/2}$$
ways to choose the bits of $x$ to flip to obtain $y$ of Hamming distance at most $\frac{m}{2} - 1$. By the entropy bound,
$${3m\choose m/2} \leq 2^{3m H(1/6)} \leq 2^{3m\frac{2}{3}} = 2^{2m}.$$
Hence 
$$|M| \geq \frac{2^{3m}}{2^{2m}} =2^m$$
as desired.
\end{proof}

@@

\begin{proof}
The lowerbound follows from the previous Theorem. It remains to give a protocol attaining the upperbound.

Let $m \in Z$ be the least integer satisfying $(6m)^2 \geq 3n$. By Lemma \ref{lemma:code} there exists a $((6m)^2, (6m)^2/3, (6m)^2/6)$-code $M'$. Choose $2^n$ elements of this code to obtain a $((6m)^2,n, (6m)^2/6))$-code $M$. Establish a bijection $\phi$ between $\{0,1\}^n$ and $M$.

Let $x \in\{0,1\}^n$ be Alice's input, and let $y \in \{0,1\}^n$ be Bob's input. Alice reinterprets $\phi(x)$ as a $6m\times 6m$ matrix, $A$, where the first row of $A$ corresponds to the first $6m$ entries of $\phi(x)$, the second row corresponds to next $6m$ entries o $\phi(x)$, etc. Alice chooses $i \in [6m]$ uniformly at random and sends $i$ and the $i$-th row of $A$, $a$ to the referee.

Analogously, Bob reinterprets $\phi(y)$ as a $6m\times 6m$ matrix $B$. Bob chooses a column $j$ of $B$ uniformly at random and send $j$ and the $j$-th row of $B$, $b$ to the referee.

The total communication cost of the protocol is thus $O(m) = O(\sqrt{n})$. The referee compares $a_j$ and $b_i$. If $a_j \neq b_i$ the referee outputs $0$. If $a_j = b_i$ then the referee outputs $1$ with probability $6/11$ and otherwise outputs $0$.

We now analyze the success probability of the algorithm. First observe that $a_j$ is $(i,j)$-th entry of $A$, and $b_i$ is the $(i,j)$-th entry of $B$. If $x=y$ then $A=B$ and the referee always enters the second case, outputting the correct answer with probability $6/11$. Consider the case where $x\neq y$. Then $A\neq B$. Since $A$ and $B$ are drawn from code $M$, they differ in at least $(6m)^2/6$ entries. Let $p$ be the probability that $A$ and $B$ differ on entry $(i,j)$. Since $(i,j)$ are chosen independently and uniformly at random, and at least $1/6$ of the total entries differ, 
$$p\geq \frac{1}{6}.$$
So the referee will output the correct answer with probability
$$p + \frac{5}{11}(1-p) \geq \frac{5}{11} + \frac{6}{11}\frac{1}{6} = \frac{6}{11}.$$
This protocol thus obtains a success probability slightly better than $1/2$, which can be boosted with a constant increase in communication cost to a protocol with success probability at least $\frac{2}{3}$
\end{proof}
